---
title: "Electronic Supplemental Material" 
subtitle: "For *Sponges presence increases diversity and abundance of fish and invertebrates in subtropical seagrass beds*"
output:
  bookdown::pdf_document2:
      toc: false
header-includes:
    \usepackage{caption}
editor_options:
  markdown:
    wrap: sentence
---

```{r setup, include=FALSE}
options(tinytex.verbose = TRUE)
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>"
)
source("scripts_comm/08_import_results.R")
if(!require(kableExtra))install.packages("kableExtra");library(kableExtra)
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

Journal name: Estuaries and Coasts

Author names: Finella M. Campanino^1^, Philina A. English^2^, Craig A. Layman^3^, Stephanie K. Archer^1^*
 
Affiliations: 

1. Louisiana Universities Marine Consortium, Chauvin, Louisiana, USA

2. Pacific Biological Station, Fisheries and Oceans Canada, Nanaimo, British Columbia, Canada

3. Center for Energy, Environment, and Sustainability, Department of Biology, Wake Forest University, Winston-Salem, North Carolina, USA
  
*corresponding author: sarcher@lumcon.edu

Submission Date: May XX, 2022


```{r expdes, echo=FALSE, fig.align='left', out.width="100%", message=FALSE, warning=FALSE}
knitr::include_graphics("figures/Experimental_design_2.pdf")
```

Figure S1. Illustrates the experimental design and sampling scheme. Data was collected prior to treatment establishment, and 1 and 12 months after treatments were established. Fish abundance and taxa richness were counted for whole plots. Other taxonomic groups and seagrass measurements were made in the locations indicated. 

\clearpage


\captionsetup[table]{labelformat=empty}

```{r fishtab}
knitr::kable(fish.models[,c(-1:-2,-7,-10)],
             # format = "latex",
             booktabs=T,
             # escape = FALSE,
    col.names = c("","k","AICc","Delta AICc",
                  "Akaike Weight","LL"),
    caption="Table S1. All models tested evaluating competing hypotheses of the drivers of fish abundance and taxa richness.")%>%
  kableExtra::kable_styling(latex_options = "scale_down",font_size=11)%>%
  kableExtra::pack_rows("Abundance",1,14)%>%
  kableExtra::pack_rows("Taxa Richness",15,28)
```

\clearpage

```{r inverttab}
knitr::kable(invert.models[,c(-1:-2,-7,-10)],
             # format = "latex",
             booktabs=T,
             # escape = FALSE,
    col.names = c("","k","AICc","Delta AICc",
                  "Akaike Weight","LL"),
    caption="Table S2. All models tested evaluating competing hypotheses of the drivers of invertebrate abundance and taxa richness.")%>%
  kable_styling(latex_options = "scale_down",font_size=11)%>%
  pack_rows("Abundance",1,14)%>%
  pack_rows("Taxa Richness",15,28)
```

